---
title: "Texas 2022 Analysis"
author: "Kirsten Sheehy"
date: "2024-05-16"
output:
  pdf_document: default
  html_document: default
---

## Overview

The following script cleans and analyzes data from Texas 2022. Fish were collected by Kirsten Sheehy and Jon Aguiñaga. Behavioral data was extracted from videos by Nishika Raghavan. Fish length was extracted from photos by Tommy (INSERT LAST NAME). Parasite data were collected by Dr. Jessica Stephenson's lab.


## Packages to Load

```{r, echo=TRUE, message=FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(tibble)
library(lubridate)
library(tidyverse)
library(ggplot2)
library(lme4)
library(pscl)
library(MASS)
library(lmerTest)
library(here)
library(knitr)
library(DHARMa)
```

## Raw Data
```{r, echo=TRUE, message=FALSE}
parasite_data <- read.csv(here("data", "copy_RAW_parasite_data_20230428.csv"))
length_data <- read.csv(here("data", "copy_Fish Length Data.csv"))
boris_data <- read.csv(here("data", "copy_RAW_Texas_BORISdata_20240515.csv"))
ID_data <- read.csv(here("data", "copy_RAW_trial_ID_data_completeonly_20240220.csv"))
```

## Tidy Data

### Parasite Data
```{r}
# rename columns to be consistent across data sets
parasite_data <- parasite_data %>% dplyr::rename(
  fish.ID = fish.id,
  site.ID = collection.site
)

# change collection.date and dissection.date to a date format (YYYY-MM-DD)
parasite_data$collection.date <- as.Date(parasite_data$collection.date,
  format = "%m/%d/%y"
)
parasite_data$dissection.date <- as.Date(parasite_data$dissection.date,
  format = "%m/%d/%y"
)

# change site names from abbreviation (WES, BR) to full (Weslaco, Brownsville)
parasite_data$site.ID <- gsub("WES", "Weslaco", parasite_data$site.ID)
parasite_data$site.ID <- gsub("BR-OP", "Brownsville", parasite_data$site.ID)

# note: the 'OP' in Brownsville stands for 'overpass'. We explored several sites
# in Brownsville, but only used the ones from the overpass for this study, so I
# simplified the name to just 'Brownsville'.
```

### Length Data

NOTE: I need to go back and check Tommy's labeling and a few measurements to be sure they're accurate.
UPDATE: There are some issues with the length data that I am working with Nishika on. For now, I'm going to omit the length data.
```{r}
# # rename columns to be consistent across data sets
# length_data <- length_data %>% dplyr::rename(
#   file.name = file_name,
#   date.image = date_image,
#   site.ID = site_ID,
#   fish.ID = fish_ID
# )
# 
# # remove spaces in fish.IDs
# length_data$fish.ID <- gsub(" ", "", length_data$fish.ID)
```


### Boris Data

First, I tidy the raw data. I rename columns and remove unnecessary ones.
```{r}
# remove unnecessary columns (largely meta data and unused features in BORIS)
boris_data <- boris_data %>% dplyr::select(
  -Observation.date, # this is just the day processed in BORIS
  -Description,
  -FPS,
  -Behavioral.category,
  -Modifiers,
  -Comment.start,
  -Comment.stop
)

# rename columns (to match up across data)
boris_data <- boris_data %>% dplyr::rename(
  pool = Subject,
  trial.length = Total.length,
  start = Start..s.,
  stop = Stop..s.,
  duration = Duration..s.
)

# split Media.file into columns to extract file name (could also use
# Observation.id, but figured this would help avoid typos made in Boris)
boris_data <- boris_data %>% tidyr::separate_wider_delim(Media.file,
  delim = "/",
  names = c(
    "file1",
    "file2",
    "file3",
    "file4",
    "file5",
    "video.ID"
  ),
  too_few = "align_end"
)

# remove the excess filepath columns
boris_data <- boris_data %>% dplyr::select(
  -file1,
  -file2,
  -file3,
  -file4,
  -file5
)

# video.id (from the file path split above) is the file name of the recording.
# It decomposes into the site ID, trial number, batch, and date recorded. The
# following code duplicates the column, then splits the information in video.id
# into separate columns.
boris_data$video.ID.split <- boris_data$video.ID
boris_data <- boris_data %>% tidyr::separate_wider_delim(video.ID.split,
  delim = "_",
  names = c(
    "site.ID",
    "trial.ID",
    "batch.ID",
    "trial.date"
  )
)

# remove the file type from the trial.date column
boris_data$trial.date <- gsub(".mov", "", boris_data$trial.date)

# change trial.date from (YYYYMMDD) to a date (YYYY-MM-DD)
boris_data$trial.date <- as.Date(boris_data$trial.date, format = "%Y%m%d")

# remove 'trial' from the data entries for trial.ID
boris_data$trial.ID <- gsub("trial", "", boris_data$trial.ID)
boris_data$trial.ID <- gsub("trail", "", boris_data$trial.ID) # had to find a few where this was a typo in the file name

# remove 'pool' from data in pool column
boris_data$pool <- gsub("Pool ", "", boris_data$pool)

# change site ID from abbreviations to full name
# note: doing them in this order is important
boris_data$site.ID <- gsub("Wes", "Weslaco", boris_data$site.ID)
boris_data$site.ID <- gsub("WES", "Weslaco", boris_data$site.ID)
boris_data$site.ID <- gsub("BR1", "Brownsville", boris_data$site.ID)
boris_data$site.ID <- gsub("BR2", "Brownsville", boris_data$site.ID)
boris_data$site.ID <- gsub("BR", "Brownsville", boris_data$site.ID)
# note: there are three entry types for Brownsville: BR, BR1, and BR2
# need to revisit lab notebook to confirm, but I believe BR1 and BR2
# are the two sides of the garage (i.e. the two cameras)
```

Now that the columns are all formatted correctly, I need to pull out the behaviors from the Behavior column into their own, separate columns.

```{r}
# start by duplicating the 'Behavior' column twice. This will be used to extract start and stop times of the three behaviors (open, hiding, startle).
boris_data <- boris_data %>%
  dplyr::mutate(behavior.start = Behavior)

boris_data <- boris_data %>%
  dplyr::mutate(behavior.stop = Behavior)

# then pivot_wider with names from behavior.start and values from start
boris_data_wide <- boris_data %>%
  tidyr::pivot_wider(
    names_from = behavior.start,
    values_from = start,
    names_prefix = "start."
  )

# do the same with stop
boris_data_wide <- boris_data_wide %>%
  tidyr::pivot_wider(
    names_from = behavior.stop,
    values_from = stop,
    names_prefix = "stop."
  )

# now, I need to get the duration of each behavior using the start and stop times
boris_data_wide <- boris_data_wide %>%
  tidyr::pivot_wider(
    names_from = Behavior,
    values_from = duration,
    names_prefix = "duration."
  )

# I'll remove stop_Startle and duration_Startle because these are 'points' not 'states' and do not have a duration
boris_data_wide <- boris_data_wide %>% dplyr::select(
  -stop.Startle,
  -duration.Startle
)
```

Now, I need to join the ID_data and boris_data_wide datasets.
```{r}
# I need a column in both ID_data and boris_data_wide to join by
# I'll create a new column that merges the file name (which already includes
# site, trial, and batch) with pool # for both data sets

boris_data_wide$merge.ID <- paste(
  boris_data_wide$video.ID,
  boris_data_wide$pool
)

ID_data$merge.ID <- paste(
  ID_data$video.ID,
  ID_data$pool
)

boris_data_merge <- boris_data_wide %>%
  left_join(ID_data, by = "merge.ID")
```


Now we tidy the merged data.
```{r}
# remove duplicate columns
boris_data_merge <- boris_data_merge %>% dplyr::select(
  -merge.ID,
  -video.ID.y,
  -pool.y,
  -site.ID.y,
  -trial.ID.y,
  -batch.ID.y,
  -trial.date.y
)

# rename columns to get rid of .x and .y appendages
boris_data_merge <- boris_data_merge %>% dplyr::rename(
  video.ID = video.ID.x,
  pool = pool.x,
  site.ID = site.ID.x,
  trial.ID = trial.ID.x,
  batch.ID = batch.ID.x,
  trial.date = trial.date.x
)

# add column for species from fish.ID
boris_data_merge <- boris_data_merge %>%
  mutate(species = fish.ID)

boris_data_merge <- boris_data_merge %>%
  separate_wider_delim(species,
    delim = "-",
    names = c(
      "species",
      "junk.num"
    )
  )

boris_data_merge <- boris_data_merge %>% dplyr::select(-junk.num)

# filling the 'start.startle' column based on fish and trial ID
boris_data_merge <- boris_data_merge %>%
  group_by(fish.ID, trial.ID) %>%
  fill(start.Startle, .direction = "downup")
```

Now, I need to standardize the trial times. When Nishika and I were observing, we sometimes recorded behaviors for longer than the prescribed 15 minutes. The following code finds the earliest behavior observation (either start.hiding or start.open) and then cuts off any observations after 15 minutes.
```{r}
# new columns with the earliest open and hiding value per fish per trial
boris_data_merge <- boris_data_merge %>%
  group_by(fish.ID, trial.ID) %>%
  mutate(
    earliest.open = min(start.Open, na.rm = TRUE),
    earliest.hiding = min(start.Hiding, na.rm = TRUE)
  )

# creates a trial cutoff time by taking the earliest behavior time (either open or closed)
# and adding 1200 seconds (20 minutes) to it
boris_data_merge <- boris_data_merge %>%
  group_by(fish.ID, trial.ID) %>%
  mutate(
    trial.end = pmin(earliest.open, earliest.hiding, na.rm = TRUE) + 1200
  )

# now, I need to remove all observations per fish per trial that exceed this cutoff time
boris_data_cutoff <- boris_data_merge %>%
  group_by(fish.ID, trial.ID) %>%
  filter(start.Open <= trial.end |
    start.Hiding <= trial.end)

# now, I need to create an 'end cap' value to replace any 'stop' behaviors
# basically, I need to close the observation (like in Boris)
# this also means I'll need to change the 'duration' columns, which are automatically
# exported from Boris.

# replace stop.Open with trial cutoff if higher than cutoff
boris_data_cutoff <- boris_data_cutoff %>%
  group_by(fish.ID, trial.ID) %>%
  mutate(stop.Open = if_else(stop.Open > trial.end, trial.end, stop.Open))

# replace stop.Open with trial cutoff if higher than cutoff
boris_data_cutoff <- boris_data_cutoff %>%
  group_by(fish.ID, trial.ID) %>%
  mutate(stop.Hiding = if_else(stop.Hiding > trial.end, trial.end, stop.Hiding))

# now, recalculate duration based on new end times
boris_data_cutoff <- boris_data_cutoff %>%
  group_by(fish.ID, trial.ID) %>%
  mutate(duration.Open = stop.Open - start.Open)

boris_data_cutoff <- boris_data_cutoff %>%
  group_by(fish.ID, trial.ID) %>%
  mutate(duration.Hiding = stop.Hiding - start.Hiding)
```

I know that I won't be using the third trial since most fish didn't get there, so I'm removing that data now.
```{r}
# removing 3rd trial since most fish didn't get three
boris_data_cutoff <- boris_data_cutoff %>%
  filter(trial.ID == "1" |
    trial.ID == "2")
```

Now let's create some columns for summary data (e.g. total time hiding)
```{r}
# time hiding per trial
total_hiding <- boris_data_cutoff %>%
  aggregate(
    duration.Hiding ~ fish.ID + trial.ID,
    sum
  )

total_hiding <- total_hiding %>%
  mutate(duplicate.fish.ID = fish.ID)

total_hiding <- total_hiding %>%
  separate_wider_delim(duplicate.fish.ID,
    delim = "-",
    names = c(
      "species",
      "junk.num"
    )
  )

total_hiding <- total_hiding %>%
  dplyr::select(-junk.num)

# time in open per trial
total_open <- boris_data_cutoff %>%
  aggregate(
    duration.Open ~ fish.ID + trial.ID,
    sum
  )

total_open <- total_open %>%
  mutate(duplicate.fish.ID = fish.ID)

total_open <- total_open %>%
  separate_wider_delim(duplicate.fish.ID,
    delim = "-",
    names = c(
      "species",
      "junk.num"
    )
  )

total_open <- total_open %>%
  dplyr::select(-junk.num)
```

Now, I want to merge total hiding and open per fish per trial.
```{r}
# I'm going to create a column with both fish ID and trial to create a unique
# row for each fish/trial combination. I'll then use this to join the hiding
# and open datasets
total_open <- total_open %>%
  unite(fish.ID.trial, c(fish.ID, trial.ID))

total_hiding <- total_hiding %>%
  unite(fish.ID.trial, c(fish.ID, trial.ID))

# merge
total_hide_open <- total_hiding %>%
  left_join(total_open, by = "fish.ID.trial")

# get rid of duplicate columns
total_hide_open <- total_hide_open %>%
  dplyr::select(-species.y)

# separate fish ID and trial
total_hide_open <- total_hide_open %>%
  separate_wider_delim(fish.ID.trial,
    delim = "_",
    names = c(
      "fish.ID",
      "trial"
    )
  )

# and for my own sanity, renaming the species column
total_hide_open <- total_hide_open %>%
  rename(species = species.x)

# time hiding and open by trial
total_hide_open <- total_hide_open %>%
  mutate(site.ID = boris_data_cutoff$site.ID[match(fish.ID, boris_data_cutoff$fish.ID)])

# hiding and open by trial, joined with parasite and size data
total_hide_open <- total_hide_open %>%
  left_join(parasite_data, by = "fish.ID", relationship = "many-to-many")

# Commenting out the length data join until I get it cleaned with Nishika.
# total_hide_open <- total_hide_open %>%
#   left_join(length_data, by = "fish.ID", relationship = "many-to-many")

total_hide_open <- total_hide_open %>%
  dplyr::select(
    -site.ID.y,
    -species.y
  )

# rename columns
total_hide_open <- total_hide_open %>% rename(
  species = species.x,
  site.ID = site.ID.x,
  total.parasites = totalpara
)

# format fish.ID
total_hide_open$fish.ID <- as.factor(total_hide_open$fish.ID)
```

We also have determined that there are a lot of errors in the data from the Brownsville site, so I'm making a subset for just Weslaco.
```{r}
ID_data_wes <- ID_data[!grepl("BR", ID_data$site.ID), ] # removing all rows with BR in the site ID (includes BR1, BR2, etc.)
total_hide_open_wes <- total_hide_open[!grepl("Brownsville", total_hide_open$site.ID), ]
parasite_data_wes <- parasite_data[!grepl("Brownsville", parasite_data$site.ID), ]
```


## Inspect data

First, some 'dummy checks' to make sure the data make sense
```{r}
# fish.IDs
unique(ID_data$fish.ID) # 68 unique IDs
unique(boris_data_cutoff$fish.ID) # 69 unique IDs -> somehow an NA was introduced?
# ok so the NA is actually 3 rows without fish.ID, so all receive NA. But all three are for the same pool and trial. I can't find a record of the ID for that pool/trial/date combo rn, so I'm going to remove. Those records are int he spiral notebook, I'll check to see if I can pin it down later.
boris_data_cutoff <- na.omit(boris_data_cutoff$fish.ID)
unique(boris_data_cutoff) # ok now we have 28 obs, just like in the ID_data
unique(parasite_data$fish.ID)
# 121, but 2 are just 'P.formosa' and 'P.latipina' because these are how fish were labeled if they didn't receive trials or if their IDs were unreadable on the labels

# Now the weslaco only datasets
unique(ID_data_wes$fish.ID) #42
unique(total_hide_open_wes$fish.ID) #42
unique(parasite_data_wes$fish.ID) #85
```

This all makes sense. The number of unique fish IDs from my notebook match up with the boris data. There are more parasite fish IDs because we sent them fish that didn't necessarily go through trials, in addition to trial fish.

```{r}
# video.ID
unique(boris_data_wide$video.ID) # 36
unique(ID_data$video.ID) # 36
```

This also makes sense. We have the same number of video IDs in the boris and ID data.

Commenting out length for now
```{r, echo=FALSE}
# # length data
# length_hist <- total_hide_open %>%
#   ggplot(mapping = aes(total_length..mm.)) +
#   geom_histogram()
# length_hist
```

Seems like a fairly normal distribution for length? possibly bimodal?

Let's take a look at the parasite data.
```{r}
# parasite data total
parasite_hist <- total_hide_open %>%
  ggplot(mapping = aes(total.parasites)) +
  geom_histogram()
parasite_hist

# parasite data weslaco
parasite_wes_hist <- total_hide_open_wes %>%
  ggplot(mapping = aes(total.parasites)) +
  geom_histogram()
parasite_wes_hist
```

Not a normal distribution. No obvious pattern here, besides a lot of zeros. Should check to see how this matches up with notes in the parasite data about specimen quality.

Let's look at the parasites by species and site.
```{r}
sp_parasite_box <- total_hide_open %>%
  ggplot(mapping = aes(
    fill = species,
    x = site.ID,
    y = total.parasites
  )) +
  geom_boxplot() +
  geom_jitter(aes(alpha = 0.5))
sp_parasite_box
```

Ok, so there is a clear pattern of way more parasites in Brownsville, generally. It also seems like there may be more parasites on amazons in both sites, but we'll see what the stats say.

Now, let's take a look at the shape of the behavior data.
```{r}
# boris_data, distributions
hiding_hist <- total_hide_open %>%
  ggplot(mapping = aes(duration.Hiding)) +
  geom_histogram()
hiding_hist

open_hist <- total_hide_open %>%
  ggplot(mapping = aes(duration.Open)) +
  geom_histogram()
open_hist

# boris_data, distributions in weslaco
hiding_wes_hist <- total_hide_open_wes %>%
  ggplot(mapping = aes(duration.Hiding)) +
  geom_histogram()
hiding_wes_hist

open_wes_hist <- total_hide_open_wes %>%
  ggplot(mapping = aes(duration.Open)) +
  geom_histogram()
open_wes_hist
```

Ok, so it looks like we have a fairly normal distribution for duration hiding, with a right skew. For time in the open, we have a floor of zero, so a strong left skew. This indicates that many fish spent the entire, or most of the trial hiding.


Now, let's take a look at the open vs. hiding data by species.
```{r}
# diff in hiding between species
species_hiding <- total_hide_open %>%
  ggplot(mapping = aes(
    x = species,
    y = duration.Hiding,
    fill = species
  )) +
  geom_boxplot()
species_hiding

# diff in open between species
species_open <- total_hide_open %>%
  ggplot(mapping = aes(
    x = species,
    y = duration.Open,
    fill = species
  )) +
  geom_boxplot()
species_open

# diff in hiding between species in weslaco
species_wes_hiding <- total_hide_open_wes %>%
  ggplot(mapping = aes(
    x = species,
    y = duration.Hiding,
    fill = species
  )) +
  geom_boxplot()
species_wes_hiding

# diff in open between species
species_wes_open <- total_hide_open_wes %>%
  ggplot(mapping = aes(
    x = species,
    y = duration.Open,
    fill = species
  )) +
  geom_boxplot()
species_wes_open
```

Again, we'll see how the stats pan out, but it looks like Amazons might spend less time hiding and more time in the open than sailfins.

## Models

### Parasites by Species

Question: is there a difference in parasite load between Amazons and Sailfins?

We'll start with a simple model.
```{r}
### code chunk is from Kate

para.lm <- lm(totalpara ~ species*site.ID, data = parasite_data)

plot(para.lm) # even just using the standard plots you can see variance heterogeneity and major non-normality

# to run all the tests in DHARMa, you first have to simulate your residuals
sim.output <- DHARMa::simulateResiduals(para.lm)

# then you can plot them 
plot(sim.output)

# gives you a bunch of tests of dispersion etc
DHARMa::testResiduals(sim.output)

# yes we can clearly see that data super zero inflated
DHARMa::testZeroInflation(sim.output)
```

Ok, so our data is zero inflated. Let's try to make some zero inflated models and check those assumptions. As Kate pointed our, DHARMa doesn't play nice with zero inflated models, so we're just going ahead without checking assumptions. I've started researching how best to test zero inflated model assumptions, but most of the packages I've found are not compatible with this version of R... more searching needed.

```{r}
mod_para_interaction <- zeroinfl(totalpara ~ species * site.ID,
  dist = "negbin",
  lin = "logit",
  data = parasite_data
)

sim.ouput <- DHARMa::simulateResiduals(mod_para_interaction) # doesn't work!
summary(mod_para_interaction)
```

KATE NOTES:

So let's explore this summary. You can see two sets of parameter estimates: 
"count model" - do our predictors affect the number of parasites we see when they are >0
"zero-inflation"  - do our predictors affect the presences of parasites

Looking at this summary there is a significant effect of site on the count of parasites (which is reassuring since we do know that the sites differ in parasite loads!)

(I have no idea what the 'log(theta) terms means... my googling suggests it's some measure of dispersion in our model but I don't feel like I have a good hand on how to interpret/if we need to interpret it)

And yeah no difference in species when only looking at weslaco. Also when only looking at brownsville too.

```{r}
# combined model
mod_para_combined <- zeroinfl(totalpara ~ species + site.ID,
  dist = "negbin",
  lin = "logit",
  data = parasite_data
)

summary(mod_para_combined) # still no effect of species


parasite_data_wes <- parasite_data %>%
  filter(site.ID == "Weslaco")

parasite_data_br <- parasite_data %>%
  filter(site.ID == "Brownsville")


# species model
mod_para_species_wes <- zeroinfl(totalpara ~ species,
  dist = "negbin",
  lin = "logit",
  data = parasite_data_wes
)

summary(mod_para_species_wes) # even within just the Weslaco site, no effect of species
hist(parasite_data_wes$totalpara)

# species model
mod_para_species_br <- zeroinfl(totalpara ~ species,
  dist = "negbin",
  lin = "logit",
  data = parasite_data_br
)

summary(mod_para_species_br) # same at Brownsville, no effect of species
hist(parasite_data_br$totalpara)

```


Let's make some plots.
```{r}
#basic
ggplot(parasite_data, aes(x = totalpara, colour = species)) +
  geom_histogram(alpha = 0.5, position = "identity")

#prettify
ggplot(parasite_data, aes(x = totalpara, fill = species, colour = species)) +
  geom_histogram(alpha = 0.3, position = "identity", binwidth = 10) + 
  scale_fill_manual(values = c("#0066FF", "#FF0033")) +
  scale_colour_manual(values = c("#0066FF", "#FF0033")) +
  xlab("Total parasite count") + 
  ylab("Frequency") + 
  theme_classic() +
  theme(legend.position = c(0.8, 0.8))

```

Zero inflated models are hungry so maybe we just create new column that bins parasites into 0/1

```{r}
parasite_data <- parasite_data %>%
  mutate(para_yes = ifelse(totalpara > 0, 1, 0))

mod_para_log <- glm(para_yes ~ species*site.ID, family = "binomial", data = parasite_data) 

sim.output <- DHARMa::simulateResiduals(mod_para_log)
plot(sim.output)

summary(mod_para_log)
```

Yeahhh.... absolutely no difference in parasite presence/absence. I think we're safe with saying that these two species are the same in parasite loads/presences etc. We can present our zero-inflated model and our logistic regression and the histogram graph as support of this.

END KATE NOTES

### Behavior and parasites

Question: Is the amount of time that fish spend hiding predicted by the number of parasites they have.

KATE NOTES

The summary is saying that we have 60 fish total but 178 observations.... I thought each fish was observed twice so how are we having more observations than 120? 

Looking through the datafile, it looks like there are some duplicate values (e.g. fish PL-41 has two trial 1's). So we're definitely going to want to carefully go through the datafile and make sure to remove these duplicates! We should do this before moving forward with digging into the models (I actually already ran a bunch of models but then realized the weird sample sizes and so deleted those until we could clear up the data file)

KIRSTEN: Those extra observations were coming from the length data. I'm still working on sorting that out with Nishika. Moving forward without length, we can use the total_hide_open_wes, which checks out with 42 fish IDs and 84 observations. I just commented out all the code that created the length data and merged it with the other data sets for now.

I'm also just going to look at the Weslaco site for now.

We'll start with a simple model.

```{r}
## Now, just with the WESLACO site ##

# filter data to just Weslaco
total_hide_open_wes <- total_hide_open[!grepl("Brownsville", total_hide_open$site.ID), ]

# behavior model
mod_hiding.lmer <- lme4::lmer(duration.Hiding ~ total.parasites*species + (1|fish.ID),
                    data = total_hide_open_wes)

plot(mod_hiding.lmer) # looks a little trumpet-y, but actually not too bad

# to run all the tests in DHARMa, you first have to simulate your residuals
beh_sim.output <- DHARMa::simulateResiduals(mod_hiding.lmer)

# then you can plot them 
plot(beh_sim.output) # also look kind of ok? Q-Q plot gets a little funky

# gives you a bunch of tests of dispersion etc
DHARMa::testResiduals(beh_sim.output)
```

I think this model's assumptions actually look ok, so I'm going to go ahead and check out our random effects and interaction. I'll keep the random effects either way.

```{r}
# the behavior model from above
mod_hiding.lmer <- lme4::lmer(duration.Hiding ~ total.parasites*species + (1|fish.ID),
                    data = total_hide_open_wes)
summary(mod_hiding.lmer)
# testing random effects
lmerTest::ranova(mod_hiding.lmer) # random effects are sig and will be included

# deciding whether to include interaction
mod_hiding_noint <- lme4::lmer(duration.Hiding ~ total.parasites + species + (1|fish.ID),
                    data = total_hide_open_wes)
summary(mod_hiding_noint)

anova(mod_hiding.lmer, mod_hiding_noint) # interaction is not sig, so I'll stick with the noint model
```


Ok so we're going to move forward with the no interaction model!

```{r}
summary(mod_hiding_noint)
```